Decentralisation of Poverty - Replication Analysis

### 2016-08-04 markdown/draft.RmD
do RCI and update markdown doc many times

---
title: "Decentralisation of Poverty - Replication Analysis"
author: "Jon Minton"
date: "2 August 2016"
output: html_document
---

@@

@@

Producing the results involved doing the following:
* Defining centroids for each place.
* Identifying the closest centroid for each datazone.
* Reweighting income deprived and total population counts from the SIMD, released at 2001 datazones, onto 2011 datazones.
* etc
* etc

@@

@@

@@

>>>>

---
title: "Decentralisation of Poverty - Replication Analysis"
author: "Jon Minton"
date: "4 August 2016"
output: html_document
---

@@

# Caveats

* Work very much in progress
* Please don't steal

@@

Producing the results involved doing the following:

* Defining centroids for each place.
* Identifying the closest centroid for each datazone.
* Reweighting income deprived and total population counts from the SIMD, released at 2001 datazones, onto 2011 datazones.
* etc
* etc

@@

### Apportioning nearest centre to datazones

@@

For Glasgow, the RCI scores (with 95% credible intervals, CrIs) for 2004 and 2012, and for the difference between these two periods, are as follows:

Quantile   | RCI, 2004 | RCI, 2012  | Change  
---------- | --------  | ---------- | -------  
median     | 0.144     | 0.093      | -0.051
2.5% CrI   | 0.142     | 0.091      | -0.054
97.5% CrI  | 0.146     | 0.095      | -0.048

From this we can conclude that there has been a statistically significant fall in RCI over these two periods using the SIMD income deprivation measure, as there had been for out-of-work benefits-based measures for 2001 and 2011 in Kavanagh at al 2016.

As with Kavanagh et al 2016 Dissimilarity index scores (D) were also calculated for both periods, along with change between them. These are shown in the table below, and also show statistically significant falls.

Quantile   | D, 2004 | D, 2012  | Change  
---------- | ------- | -------- | ------- 
median     | 0.368   | 0.315    | -0.053
2.5% CrI   | 0.366   | 0.313    | -0.056
97.5% CrI  | 0.370   | 0.317    | -0.051

@@

## Plotting temporal dependence

The following plot shows scatterplots of the ID proportions in each Scottish datazone in 2004, 2009 and 2012. Within these plots the blue line indicates parity between values on the x and y axis, and the red dashed line shows a nonlinear smoother function between the two ID proportions being plotted, and hence the level of deviation between the actual relationship over the two years and parity. 
```{r show_temporal_dep, echo = FALSE, warning = FALSE, EVAL = TRUE}

simd_2011_reweighted %>% 
  filter(year %in% c(2004, 2012)) %>% 
  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
  select(dz_2011, year, prop_id) %>% 
  spread(year, prop_id) %>% 
  rename(prop_id_2004 = `2004`, prop_id_2012 = `2012`) %>% 
  ggplot(., aes(x = prop_id_2004, y = prop_id_2012)) + 
  geom_point(alpha = 0.1) + 
  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
  stat_smooth(colour = "red", linetype = "dashed") + 
  labs(
    x = "ID prop, 2004", y = "ID prop, 2012",
    title = "2012 against 2004"
       ) -> s_04_12

simd_2011_reweighted %>% 
  filter(year %in% c(2004, 2006)) %>% 
  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
  select(dz_2011, year, prop_id) %>% 
  spread(year, prop_id) %>% 
  rename(prop_id_2004 = `2004`, prop_id_2006 = `2006`) %>% 
  ggplot(., aes(x = prop_id_2004, y = prop_id_2006)) + 
  geom_point(alpha = 0.1) + 
  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
  stat_smooth(colour = "red", linetype = "dashed") + 
  labs(
    x = "ID prop, 2004", y = "ID prop, 2006",
    title = "2006 against 2004"
       ) -> s_04_06

simd_2011_reweighted %>% 
  filter(year %in% c(2006, 2009)) %>% 
  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
  select(dz_2011, year, prop_id) %>% 
  spread(year, prop_id) %>% 
  rename(prop_id_2006 = `2006`, prop_id_2009 = `2009`) %>% 
  ggplot(., aes(x = prop_id_2006, y = prop_id_2009)) + 
  geom_point(alpha = 0.1) + 
  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
  stat_smooth(colour = "red", linetype = "dashed") + 
  labs(
    x = "ID prop, 2006", y = "ID prop, 2009",
    title = "2009 against 2006"
    ) -> s_06_09

simd_2011_reweighted %>% 
  filter(year %in% c(2009, 2012)) %>% 
  mutate(prop_id = pop_incomedeprived / pop_total) %>% 
  select(dz_2011, year, prop_id) %>% 
  spread(year, prop_id) %>% 
  rename(prop_id_2009 = `2009`, prop_id_2012 = `2012`) %>% 
  ggplot(., aes(x = prop_id_2009, y = prop_id_2012)) + 
  geom_point(alpha = 0.1) + 
  geom_abline(colour = "blue", slope = 1, intercept = 0) + 
  stat_smooth(colour = "red", linetype = "dashed") + 
  labs(
    x = "ID prop, 2009", y = "ID prop, 2012",
    title = "2012 against 2009"
  ) -> s_09_12

plot_grid(s_04_12, s_04_06, s_06_09, s_09_12, nrow = 2)

There were indications in the maps above that the greatest changes occurred between the 2009 and 2012 SIMD releases, suggesting that the temporal association between 2004 and 2009 ID proportions would likely be strong compared with the temporal association between 2009 and 2012.

The first of the plots, 2004 compared with 2012, shows strong spatial dependence in areas with low to moderate ID proportions, but much less temporal dependence in areas with high ID proportions in 2004. The second subfigure plots 2004 against 2006 values, and shows a near linear temporal dependence for the complete range of 2004 values, with a global trend towards lower ID proportions throughout Scotland.  The temporal relationship between 2006 and 2009 starts to become nonlinear at the upper range of 2006 values, and this nonlinearity at the upper range becomes much stronger between 2009 and 2012.

These plots indicate a complex change in the temporal dependence of ID proportions, likely occurring over the period 2008 to 2012. SIMD 2016 values will be invaluable in indicating whether these changes have continued, altered or settled between 2012 and 2016.
